---
title: "National Travel Trends"
author: "2001, 2009, 2017 National Household Travel Survey, Combined Estimates"
output: html_document
---

<link rel="stylesheet" href="//code.jquery.com/ui/1.12.1/themes/base/jquery-ui.css">
<script src="https://code.jquery.com/ui/1.12.1/jquery-ui.js"></script>
<script src="https://omnipotent.net/jquery.sparkline/2.1.2/jquery.sparkline.min.js"></script>


```{r setup, echo = FALSE, eval = TRUE, message = FALSE, warning = FALSE}

source("process-estimates.R")
source("create-trend-table.R")

estimate_html_content <- list()
trend_tables <- list()
for(estimate in 1:nrow(estimates)) {

	html_content <- list(
			"2001" = make_chart(results_2001[[estimate]]),
			"2009" = make_chart(results_2009[[estimate]]),
			"2017" = make_chart(results_2017[[estimate]])
	)
	
	estimate_title <- paste(estimates[estimate, "title"])
	estimate_html_content[[estimate_title]] <- html_content
	
	trend_tables[[estimate]] <- create_trend_table(estimate)

}

# include estimates.csv as documentation
file_size <- file.info("estimates.csv")[, "size"]
pdf_base64_string <- RCurl::base64Encode(readBin("estimates.csv", what="raw", n=file_size+16)) # need +16 byte overflow buffer
csv_download_link <- htmltools::a(href=paste0("data:application/octet-stream;base64,", pdf_base64_string), download="estimates.csv", "estimates.csv")

```

# {.tabset .tabset-dropdown}

## <i>Click here to select an estimate</i>
<div class="alert alert-success alert-dismissible"">
<a href="#" class="close" data-dismiss="alert" aria-label="close" style="transform:scale(2);">&times;</a>

* The `r paste(nrow(estimates))` multi-year estimates in this report were generated from publicly available [National Household Travel Survey data](https://nhts.ornl.gov)

* Estimates and visuals created using [summarizeNHTS](https://github.com/Westat-Transportation/summarizeNHTS)

* Specification for estimates can be reviewed in `r paste(csv_download_link)`

</div>

<!-- loop through estimate records and create pages using rmarkdown syntax -->
```{r, echo = FALSE, eval = TRUE}

out <- lapply(seq_along(estimate_html_content), function(i) {

	paste(
	  knitr::knit_expand(text = sprintf("## %s\n", names(estimate_html_content)[i])),
	  knitr::knit_expand(text = "\n<div class='row'>"),
	  knitr::knit_expand(text = "\n<div class='col-md-4'>"),
	  knitr::knit_expand(text = "\n### 2001"),
	  knitr::knit_expand(text = "\n```{r echo=FALSE}"),
	  knitr::knit_expand(text = sprintf("\nestimate_html_content[[%d]][['2001']]", i)),
		knitr::knit_expand(text = "\n```"),
		knitr::knit_expand(text = "\n</div>"),
		knitr::knit_expand(text = "\n<div class='col-md-4'>"),
	  knitr::knit_expand(text = "\n### 2009"),
		knitr::knit_expand(text = "\n```{r echo=FALSE}"),
	  knitr::knit_expand(text = sprintf("\nestimate_html_content[[%d]][['2009']]", i)),
	  knitr::knit_expand(text = "\n```"),
	  knitr::knit_expand(text = "\n</div>"),
		knitr::knit_expand(text = "\n<div class='col-md-4'>"),
	  knitr::knit_expand(text = "\n### 2017"),
		knitr::knit_expand(text = "\n```{r echo=FALSE}"),
	  knitr::knit_expand(text = sprintf("\nestimate_html_content[[%d]][['2017']]", i)),
	  knitr::knit_expand(text = "\n```"),
	  knitr::knit_expand(text = "\n</div>"),
	  knitr::knit_expand(text = "\n</div>"),
		knitr::knit_expand(text = "\n<div class='row'>"),
		knitr::knit_expand(text = "\n<div class='col-md-12'>"),
		knitr::knit_expand(text = "\n### Trend Significance Table"),
		knitr::knit_expand(text = "\n```{r echo=FALSE}"),
		knitr::knit_expand(text = sprintf("\ntrend_tables[[%d]]", i)),
		knitr::knit_expand(text = "\n```"),
		knitr::knit_expand(text = "\n</div>"),
		knitr::knit_expand(text = "\n</div>"),
	collapse = '\n')

})

```

`r paste(knitr::knit(text = paste(out, collapse = '\n')))`

<!-- ------------------------------------------------- -->
<!-- css -->
<style>

	body {
		background-color: #d9e5f2;
	}
	
	.main-container {
		background-color: #ffffff;
		height: 100vh;
	}
	
	.tabset-dropdown > .nav-tabs {
		width: 100%;
		display: inline-block;
		overflow: hidden;
	}
	li[role=presentation] > a[role=tab], .nav-tabs:not(.nav-tabs-open) {
		cursor: pointer !important;
	}
	li[role=presentation]:not(.active) > a[role=tab] {
	 width: 100%;
	}

</style>

<!-- sparklines tooltip -->
<script> 
	$(function() {
	    $(".gmisc_table").tooltip({
	      content:function(){
	        var regex = new RegExp("\\|", "g");
	        var title_tag = this.getAttribute("title");
	        var title_tag = title_tag.replace(regex, "<br>");
	        return '<div> ' + title_tag + '</div>';
	      }
	  });
	});
</script>

<!-- sparklines -->
<script type="text/javascript">
		$('.sparklines').sparkline('html', {
		  disableTooltips: true,
		  type: 'line', 
		  chartRangeMin: '-1', 
		  chartRangeMax: '1', 
		  width: '35px',
		  height: '30px',
		  spotColor: 'black',
		  minSpotColor: 'black',
		  maxSpotColor: 'black',
		  valueSpots: {':100': 'black'}
		});
</script>

<!-- <div id="dialog_container"></div> -->
<!-- <script> -->
<!-- 	$(function() { -->
<!-- 		$("#dialog_container").dialog({ -->
<!-- 			autoOpen: false,   -->
<!-- 		}); -->

<!-- 		function myHandler(event) { -->
<!-- 			console.log(event.data.param1); -->
<!-- 			$("#dialog_container").replaceWith(""); -->
<!-- 			$("#dialog_container").append($(event.data.param1).contents()); -->
<!-- 			$("#dialog_container").dialog("open"); -->
<!-- 		} -->

<!-- 		$(".col-md-4 > h3").each(function() { -->
<!-- 			thename = $(this).attr("name"); -->
<!-- 			$(this).on("click",{param1:thename}, myHandler); -->
<!-- 		}); -->

<!-- 	}); -->
<!-- </script> -->

<!-- trend significance table column widths -->
<script>
	$(".gmisc_table tr:not(:first-child) th:odd").css({"max-width": "75px", "min-width": "75px"})
	$(".gmisc_table tr td:first-child").css({"min-width": "150px"})
</script>
